---
title: "STAT 331 Portfolio"
author: "Karthik Balaji"
format: html 
embed-resources: true
layout: margin-left
editor: visual
execute: 
  eval: false
  echo: true
---

[**My Grade:**]{.underline} I believe my grade equivalent to course work evidenced below to be an \_\_.

[**Learning Objective Evidence:**]{.underline} In the code chunks below, provide code from Lab or Challenge assignments where you believe you have demonstrated proficiency with the specified learning target. Be sure to specify **where** the code came from (e.g., Lab 4 Question 2).

## Working with Data

**WD-1: I can import data from a *variety* of formats (e.g., csv, xlsx, txt, etc.).**

-   `csv` Example 1

```{r}
#| label: wd-1-csv-1

# Lab 3 Question 2

teacher_evals <- read_csv(here::here("Week3", "Lab3", "teacher_evals.csv"))
```

-   `csv` Example 2

```{r}
#| label: wd-1-csv-2

# Lab 4 Question 0
childcare_costs <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-09/childcare_costs.csv')

counties <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-09/counties.csv')

tax_rev <- read_csv('https://atheobold.github.io/groupworthy-data-science/labs/instructions/data/ca_tax_revenue.csv')

```

-   `xlsx`

```{r}
#| label: wd-1-xlsx

# Haven't used apart from PA and Check-Ins
```

**WD-2: I can select necessary columns from a dataset.**

-   Example selecting specified columns

```{r}
#| label: wd-2-ex-1

# Lab 3 Question 5
teacher_evals_clean <- teacher_evals |>
    rename(sex = gender) |>
    filter(no_participants >= 10) |>
    mutate(
      across(c(teacher_id, course_id), as.character),
      across(c(question_no, no_participants, seniority), as.integer)
    ) |>
    select(
      course_id,
      teacher_id,
      question_no,
      no_participants,
      resp_share,
      SET_score_avg,
      percent_failed_cur,
      academic_degree,
      seniority,
      sex
    )
```

-   Example removing specified columns

```{r}
#| label: wd-2-ex-2

# Modified Lab 5 Step 1 to use removal of irrelevant columns
crime_scene_report |>
  mutate(date = ymd(date)) |>
  filter(date == mdy("January 15, 2018"),
         type == "murder",
         city == "SQL City") |>
  select(-date, -type, -city)
```

-   Example selecting columns based on logical values (e.g., `starts_with()`, `ends_with()`, `contains()`, `where()`)

```{r}
#| label: wd-2-ex-3

#| label: recreate-plot
#| fig-width: 10
#| fig-height: 3.3

# Modified Lab 4 Question 7 to use starts_with instead of rename, makes code a lot cleaner
ca_childcare |>
  pivot_longer(
    cols = starts_with("mc_"),
    names_to  = "age_group",
    names_prefix = "mc_",
    values_to = "median_price"
  ) |>
  mutate(
    age_group = recode(age_group,
                       infant = "Infant",
                       toddler = "Toddler",
                       preschool = "Preschool"),
    region = fct_reorder2(region, study_year, median_price),
    age_group = fct_relevel(age_group, "Infant", "Toddler", "Preschool")
  ) |>
  ggplot(mapping = aes(x = study_year, y = median_price, color = region)) +
    geom_point() +
    scale_x_continuous(breaks = seq(2008, 2018, 2)) +
    scale_y_continuous(breaks = seq(0, 500, 100),
                       limits = c(100, 500)) +
    geom_smooth(method = "loess") +
    theme_bw() +
    labs(title = "Weekly Median Price for Center-Based Childcare ($)",
         x = "Study Year",
         y = NULL,
         color = "California Region") 
```

**WD-3: I can filter rows from a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   Numeric Example 1

```{r}
#| label: wd-3-numeric-ex-1

# Lab 4 Question 5
ca_childcare |> 
  filter(study_year %in% c(2008, 2018)) |>
  group_by(region, study_year) |>
  summarize(median_income = median(mhi_2018)) |>
  pivot_wider(names_from = study_year,
              values_from = median_income) |>
  arrange(desc(`2018`))
```

-   Numeric Example 2

```{r}
#| label: wd-3-numeric-ex-1

#Lab 3 Question 5
teacher_evals_clean <- teacher_evals |>
    rename(sex = gender) |>
    filter(no_participants >= 10) |>
    mutate(
      across(c(teacher_id, course_id), as.character),
      across(c(question_no, no_participants, seniority), as.integer)
    ) |>
    select(
      course_id,
      teacher_id,
      question_no,
      no_participants,
      resp_share,
      SET_score_avg,
      percent_failed_cur,
      academic_degree,
      seniority,
      sex
    )
```

-   Character Example 1 (any context)

```{r}
#| label: wd-3-character

# Lab 4 Question 2
ca_childcare <- left_join(x = childcare_costs,
                          y = counties,
                          by = "county_fips_code") |>
                filter(state_abbreviation == "CA") 
```

-   Character Example 2 (example must use functions from **stringr**)

```{r}
#| label: wd-3-string

# Lab 5 Part 2
person |>
  filter(str_detect(name, pattern = "^Annabel"),
         address_street_name == "Franklin Ave") |>
  left_join(interview, by = join_by(id == person_id)) |>
  select(transcript)
```

-   Date (example must use functions from **lubridate**)

```{r}
#| label: wd-3-date

# Lab 5 Part 4
facebook_event_checkin |>
  mutate(date = ymd(date)) |>
  filter(event_name == "SQL Symphony Concert",
         year(date) == 2017,
         month(date) == 12) |>
  count(person_id, name = "attendances") |>   
  filter(attendances == 3) |>
  left_join(person, by = join_by(person_id == id)) |>
  left_join(drivers_license, by = join_by(license_id == id)) |>
  filter(gender == "female",
         hair_color == "red",
         height %in% c(65, 66, 67),
         car_make == "Tesla",
         car_model == "Model S") |>
  select(name)
```

**WD-4: I can modify existing variables and create new variables in a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   Numeric Example 1

```{r}
#| label: wd-4-numeric-ex-1

```

-   Numeric Example 2

```{r}
#| label: wd-4-numeric-ex-2

```

-   Factor Example 1 (renaming levels)

```{r}
#| label: wd-4-factor-ex-1

```

-   Factor Example 2 (reordering levels)

```{r}
#| label: wd-4-factor-ex-2

```

-   Character (example must use functions from **stringr**)

```{r}
#| label: wd-4-string

```

-   Date (example must use functions from **lubridate**)

```{r}
#| label: wd-4-date

```

**WD-5: I can use mutating joins to combine multiple dataframes.**

-   `left_join()` Example 1

```{r}
#| label: wd-5-left-ex-1

```

-   `right_join()` Example 1

```{r}
#| label: wd-5-right

```

-   `left_join()` **or** `right_join()` Example 2

```{r}
#| label: wd-5-left-right-ex-2

```

-   `inner_join()` Example 1

```{r}
#| label: wd-5-inner-ex-1

```

-   `inner_join()` Example 2

```{r}
#| label: wd-5-inner-ex-2

```

**WD-6: I can use filtering joins to filter rows from a dataframe.**

-   `semi_join()`

```{r}
#| label: wd-6-semi

```

-   `anti_join()`

```{r}
#| label: wd-6-anti

```

**WD-7: I can pivot dataframes from long to wide and visa versa**

-   `pivot_longer()`

```{r}
#| label: wd-7-long

```

-   `pivot_wider()`

```{r}
#| label: wd-7-wide

```

## Reproducibility

**R-1: I can create professional looking, reproducible analyses using RStudio projects, Quarto documents, and the here package.**

The following assignments satisfy the above criteria:

-   Example 1
-   Example 2
-   Example 3
-   Example 4
-   Example 5

**R-2: I can write well documented and tidy code.**

-   Example of **ggplot2** plotting

```{r}
#| label: r-2-1

```

-   Example of **dplyr** pipeline

```{r}
#| label: r-2-2

```

-   Example of function formatting

```{r}
#| label: r-2-3

```

**R-3: I can write robust programs that are resistant to changes in inputs.**

-   Example (any context)

```{r}
#| label: r-3-example

```

-   Example (function stops)

```{r}
#| label: r-3-function-stops

```

## Data Visualization & Summarization

**DVS-1: I can create visualizations for a *variety* of variable types (e.g., numeric, character, factor, date)**

-   At least two numeric variables

```{r}
#| label: dvs-1-num

```

-   At least one numeric variable and one categorical variable

```{r}
#| label: dvs-2-num-cat

```

-   At least two categorical variables

```{r}
#| label: dvs-2-cat

```

-   Dates (time series plot)

```{r}
#| label: dvs-2-date

```

**DVS-2: I use plot modifications to make my visualization clear to the reader.**

-   I can modify my plot theme to be more readable

```{r}
#| label: dvs-2-ex-1

```

-   I can modify my colors to be accessible to anyone's eyes

```{r}
#| label: dvs-2-ex-2

```

-   I can modify my plot titles to clearly communicate the data context

```{r}
#| label: dvs-2-ex-3

```

-   I can modify the text in my plot to be more readable

```{r}
#| label: dvs-2-ex-4

```

-   I can reorder my legend to align with the colors in my plot

```{r}
#| label: dvs-2-ex-5

```

**DVS-3: I show creativity in my visualizations**

-   I can use non-standard colors (Example 1)

```{r}
#| label: dvs-3-1-ex-1

```

-   I can use non-standard colors (Example 2)

```{r}
#| label: dvs-3-1-ex-2

```

-   I can use annotations (e.g., `geom_text()`)

```{r}
#| label: dvs-3-2

```

-   I can choose creative geometries (e.g., `geom_segment()`, `geom_ribbon)()`)

```{r}
#| label: dvs-3-3

```

**DVS-4: I can calculate numerical summaries of variables.**

-   Example using `summarize()`

```{r}
#| label: dvs-4-summarize

```

-   Example using `across()`

```{r}
#| label: dvs-4-across

```

**DVS-5: I can find summaries of variables across multiple groups.**

-   Example 1

```{r}
#| label: dvs-5-1

```

-   Example 2

```{r}
#| label: dvs-5-2

```

**DVS-6: I can create tables which make my summaries clear to the reader.**

-   I can modify my column names to clearly communicate the data context

```{r}
#| label: dvs-6-ex-1

```

-   I can modify the text in my table to be more readable (e.g., bold face for column headers)

```{r}
#| label: dvs-6-ex-2

```

-   I can arrange my table to have an intuitive ordering

```{r}
#| label: dvs-6-ex-3

```

**DVS-7: I show creativity in my tables.**

-   I can use non-default colors

```{r}
#| label: dvs-7-ex-1

```

-   I can modify the layout of my table to be more readable (e.g., `pivot_longer()` or `pivot_wider()`)

```{r}
#| label: dvs-7-ex-2

```

## Program Efficiency

**PE-1: I can write concise code which does not repeat itself.**

-   using a single function call with multiple inputs (rather than multiple function calls)

```{r}
#| label: pe-1-one-call

```

-   using `across()`

```{r}
#| label: pe-1-across

```

-   using functions from the `map()` family

```{r}
#| label: pe-1-map-1

```

**PE-2: I can write functions to reduce repetition in my code.**

-   Example 1: Function that operates on vectors

```{r}
#| label: pe-2-1

```

-   Example 2: Function that operates on data frames

```{r}
#| label: pe-2-2

```

-   Example 3: Function that operates on vectors *or* data frames

```{r}
#| label: pe-2-3

```

**PE-3:I can use iteration to reduce repetition in my code.**

-   using `across()`

```{r}
#| label: pe-3-across

```

-   using a `map()` function with **one** input (e.g., `map()`, `map_chr()`, `map_dbl()`, etc.)

```{r}
#| label: pe-3-map-1

```

-   using a `map()` function with **more than one** input (e.g., `map_2()` or `pmap()`)

```{r}
#| label: pe-3-map-2

```

**PE-4: I can use modern tools when carrying out my analysis.**

-   I can use functions which are not superseded or deprecated

```{r}
#| label: pe-4-1

```

-   I can connect a data wrangling pipeline into a `ggplot()`

```{r}
#| label: pe-4-2

```

## Data Simulation & Statisical Models

**DSSM-1: I can simulate data from a *variety* of probability models.**

-   Example 1

```{r}
#| label: dsm-1-1

```

-   Example 2

```{r}
#| label: dsm-1-2

```

**DSSM-2: I can conduct common statistical analyses in R.**

-   Example 1

```{r}
#| label: dsm-2-1

```

-   Example 2

```{r}
#| label: dsm-2-2

```

-   Example 3

```{r}
#| label: dsm-2-3

```

## Revising My Thinking

<!-- How did you revise your thinking throughout the course? How did you revise your thinking on the code examples you have provided in your portfolio? -->

<!-- For the revisions included in your Portfolio, to help me understand the nature of your revisions, please denote somehow the feedback I provided you (e.g., boldface, italics, colored text) before your revisions. -->

## Extending My Thinking

<!-- How did you extended your thinking throughout the course? How did you extend your thinking on the code examples you have provided in your portfolio? -->

## Peer Support & Collaboration

<!-- Include an image or a description of feedback you gave that you are proud of (either in a peer review or in Discord). -->

<!-- Include a description of how you grew as a collaborator through the weekly pair programming activities.   -->
